A supertree pipeline for summarizing phylogenetic and taxonomic information for millions of species

### 2017-01-19 doc/paper1-submission/paper5.tex
fig legends and referencing

These optimality criteria help to define what it means for the supertree
to represent the input trees, as well as justifying and explaining
why various features of the supertree exist. The pipeline described
below produces a supertree that satisfies the first three optimality
criteria and is a greedy approximation of a solution to the fourth
goal. It is not guaranteed to display as many input nodes as possible.
Even if the summary tree does accomplish goal 4, it is not necessarily
a \emph{unique} optimum. The pipeline takes a greedy approach to producing
a summary tree by attempting to add groupings from the trees in order
of the trees ranking. This can be viewed as a greedy solution to the
problem of finding the tree with the maximum sum of displayed groups'
weighted scores criterion (MSDGWS, described in the appendix A) where the
weights from the trees are so extreme that displaying one group from
a highly ranked tree is preferred to displaying all of the groupings
from lower ranked trees.

@@

So, we also consider MRP and MinCut.  MRP can display edges not found
in any input tree, and so fails criterion \#1.  MRP is used in many
studies that attempt to infer the true tree from a lot of input
trees.  To some extent, MRP tries to treat each tree as data, which
differs from our approach here.

MinCut (and Modified MinCut) attempts to favor relationships present
in multiple input trees during conflict resolution.

Both MRP and MinCut attempt to adjudicate conflicts in a way that
favors relationships present in more input trees.  Thus, they perform
something more like a weighted average.  However, the reason the
supertree resolves conflicts in a particular way is not clear and
transparent, and does not allow a method for curators to adjust the
supertree.

@@

To produce a tree that spans all of the taxa sampled in the exemplified
set of input trees, we graft the subproblem solutions into a single
tree (stored as the \texttt{grafted\_solution/grafted\_solution.tre}
by propinquity). Recall that each non-root uncontested taxon used
for decomposition occurs as a leaf taxon in one subproblem and as
a root taxon in one other subproblem. Thus, the grafting operation
simply consists of reading all of the subproblem solutions into memory
and then merging the nodes that are labeled with the same OTT ID.

\begin{figure}
\begin{centering}
\includegraphics[scale=0.5]{Figures__grafted_solution__grafted_solution_ottnames\lyxdot tre}
\par\end{centering}
\caption{Grafted solution. The nodes E and F are monotypic here, but will end
up being polytypic after unpruning is performed.}
\end{figure}

>>>>

These optimality criteria help to define what it means for the supertree
to represent the input trees, as well as justifying and explaining
why various features of the supertree exist. The pipeline described
below produces a supertree that satisfies the first three optimality
criteria and is a greedy approximation of a solution to the fourth
goal. It is not guaranteed to display as many input nodes as possible.
Even if the summary tree does accomplish goal 4, it is not necessarily
a \emph{unique} optimum. The pipeline takes a greedy approach to producing
a summary tree by attempting to add groupings from the trees in order
of the trees ranking. This can be viewed as a greedy solution to the
problem of finding the tree with the maximum sum of displayed groups'
weighted scores criterion (MSDGWS, described in the appendix A) where the
weights from the trees are so extreme that displaying one group from
a highly ranked tree is preferred to displaying all of the groupings
from lower ranked trees.
Note that in this extreme form of weighting, we do not have to calculate
the score of a tree, we simply traverse the groupings by order of rank
and add a group if it is compatible with the previously added groups.

@@

We also consider matrix representation with parsimony (MRP) and MinCut.
MRP can display edges not found
in any input tree, and so fails criterion \#1.  MRP is used in many
studies that attempt to infer the true tree from a lot of input
trees.  To some extent, MRP tries to treat each tree as data, which
differs from our approach here.
MinCut (and Modified MinCut) attempts to favor relationships present
in multiple input trees during conflict resolution.
Both MRP and MinCut attempt to adjudicate conflicts in a way that
favors relationships present in more input trees.  Thus, they perform
something more like a weighted average.  However, the reason the
supertree resolves conflicts in a particular way is not clear and
transparent, and does not allow a method for curators to adjust the
supertree.

@@

To produce a tree that spans all of the taxa sampled in the exemplified
set of input trees, we graft the subproblem solutions into a single
tree (stored as the \texttt{grafted\_solution/grafted\_solution.tre}
by propinquity); see figure \ref{fig:grafted-solution}. Recall that each non-root uncontested taxon used
for decomposition occurs as a leaf taxon in one subproblem and as
a root taxon in one other subproblem. Thus, the grafting operation
simply consists of reading all of the subproblem solutions into memory
and then merging the nodes that are labeled with the same OTT ID.

\begin{figure}
\begin{centering}
\includegraphics[scale=0.5]{Figures__grafted_solution__grafted_solution_ottnames\lyxdot tre}
\par\end{centering}
\caption{Grafted solution produced from the subproblems from
figure \ref{fig:subproblem-solutions.} and which is the 
backbone onto which taxa that are not included in any phylogeny
will be placed.}
\label{fig:grafted-solution}
\end{figure}
